---
title: "Emissions & Preferences"
author: "Martin Blasi, Jonas Wallstein & Carles Manchado"
date: "2022-11-28"
header-includes:
  -\newcommand{\bx}{{\bf x}}
  -\newcommand{\by}{{\bf y}}
  -\newcommand{\bv}{{\bf v}}
  -\newcommand{\be}{{\bf e}}
  -\newcommand{\bu}{{\bf u}}
output: 
  slidy_presentation:
    incremental: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Emissions and Preferences

- CO2 emissions: Total and per sector (9) from NACE (Statistical classification of economic activities in the European Community) in tonnes of CO2 per capita

- Together with GDP per capita from OECD

- Preference parameters from Global Preference Survey, Falck et al. (2012)

- n=74 countries

- p=17 variables (subset with p=8)

- Methods: PCA & Clustering

- Goal: Hidden interesting patterns behind pollution behaviour and fundamental preference paramters





## Results

- PCAs deliver poor picture

- First two PCs only explain 46% of the variance

- Main reason: Correlation between preference paramters and emissions is not too great

- Clustering delivers four/five groups that sort of make sense






## Data description

Load the data
```{r}
load("C:/Users/blasi/OneDrive/Desktop/UPF Studies/multivariate data analysis/Class/Group project/1 LOCAL COPIES/gps/Emissions_and_Preferences.RData")
```
Subset
```{r}
library("dplyr")
df.empref1 =
  df.empref %>%
    select(-c(country))

rownames(df.empref1) <- df.empref1[,1]

df.empref2 =
  df.empref1 %>%
    select(-c(iso_code))

#subset of preferences and only TOTAL emissions
df.empref2sub =
    df.empref2 %>%
      select(-c(emissions.agriculture, emissions.buildings, emissions.extraction, emissions.manufacturing, emissions.maritime, emissions.oil.and.gas, emissions.power, emissions.transport ,emissions.waste))

head(df.empref2)
```
Scale and covariance matrix
```{r}
X <- as.matrix(df.empref2)

Xs <- scale(X)
R <- cov(Xs)
R
```





## PCA
```{r}
l <- eigen(R)$values
V <- eigen(R)$vectors
Y <- Xs %*% V
# the portion of the variance that retains each PC
R2 = l / sum(l) *100
# accumulating this quantities:
cumsum( l / sum(l))

#automatically
library(ade4)
library(factoextra)
pca.empref = dudi.pca(df.empref2, scannf = FALSE)
pca.empref$eig
cumsum(pca.empref$eig)/sum(pca.empref$eig)
fviz_eig(pca.empref, geom = "bar", bar_width = 0.3, addlabels=T) + ggtitle("")
```

```{r}
#Plot
plot(Y[,1:2],
  type="n", # don't plot anything now
  asp = 1, # aspect ratio, very important!
  main=paste("Variables plot, expl. var is ",
               round(R2[1]+R2[2],1), " %"),
  xlab=paste("CP1(", round(R2[1],1), "%)"), # title x axis
  ylab=paste("CP2(", round(R2[2],1), "%)") # title y axis
)

# add the axis
abline(v=0, col="blue")
abline(h=0, col="blue")

# the dots and names
points(Y[,1:2], pch=".") # plot a dot per row
text(Y[,1:2],row.names(df.empref2),cex=0.6)
```




## SVD and Variables Plot
```{r}
#Carry out the Singular Value Decomposition
svd <- svd(Xs)


Utilde <- svd$u[,1:2] %*% sqrt(diag(svd$d[1:2]))
Vtilde <- svd$v[,1:2] %*% sqrt(diag(svd$d[1:2]))


nn <- paste(rownames(Xs))
xlim <- ylim <- range(c(range(1.5*Vtilde)))
plot(NULL,xlab='First biplot component',ylab='Second biplot component',xlim=xlim,ylim=ylim,cex.lab=0.5)
arrows(x0=0,y0=0,x1=Vtilde[,1],y1=Vtilde[,2],col='red', code=2, length=0.1)
text(.9*Vtilde[,1],Vtilde[,2],paste(colnames(Xs)),pos=c(1,3),col='red',cex=0.5)
```

Biplot
```{r}
nn <- paste(rownames(Xs))
xlim <- ylim <- range(c(range(Vtilde),range(1.5*Vtilde)))
plot(Utilde,xlab='First biplot component',ylab='Second biplot component',xlim=xlim,ylim=ylim,cex.lab=1.25)
text(Utilde[,1:2],nn,pos=c(4,1,2),cex=0.5)
arrows(x0=0,y0=0,x1=Vtilde[,1],y1=Vtilde[,2],col='red', code=2, length=0.1)
text(.9*Vtilde[,1],Vtilde[,2],paste(colnames(Xs)),pos=c(1,3),col='red',cex=0.5)

```




## Subset Analysis

Total emissions only

Scale and covariance matrix
```{r}
Xsub <- as.matrix(df.empref2sub)

Xssub <- scale(Xsub)
Rs <- cov(Xssub)
```
PCA
```{r}
ls <- eigen(Rs)$values
Vs <- eigen(Rs)$vectors
Ys <- Xssub %*% Vs
# the portion of the variance that retains each PC
R2s = ls / sum(ls) *100
# accumulating this quantities:
cumsum( ls / sum(ls))

#automatically
library(ade4)
library(factoextra)
pca.emprefsub = dudi.pca(df.empref2sub, scannf = FALSE)
pca.emprefsub$eig
cumsum(pca.emprefsub$eig)/sum(pca.emprefsub$eig)
fviz_eig(pca.emprefsub, geom = "bar", bar_width = 0.3, addlabels=T) + ggtitle("")
```

```{r}
#Plot
plot(Ys[,1:2],
  type="n", # don't plot anything now
  asp = 1, # aspect ratio, very important!
  main=paste("Variables plot, expl. var is ",
               round(R2s[1]+R2s[2],1), " %"),
  xlab=paste("CP1(", round(R2s[1],1), "%)"), # title x axis
  ylab=paste("CP2(", round(R2s[2],1), "%)") # title y axis
)

# add the axis
abline(v=0, col="blue")
abline(h=0, col="blue")

# the dots and names
points(Ys[,1:2], pch=".") # plot a dot per row
text(Ys[,1:2],row.names(df.empref2sub),cex=0.6)
```




## SVD and Variables Plot
```{r}
#Carry out the Singular Value Decomposition
svdsub <- svd(Xssub)


Utildes <- svdsub$u[,1:2] %*% sqrt(diag(svdsub$d[1:2]))
Vtildes <- svdsub$v[,1:2] %*% sqrt(diag(svdsub$d[1:2]))


nn <- paste(rownames(Xssub))
xlim <- ylim <- range(c(range(1.5*Vtildes)))
plot(NULL,xlab='First biplot component',ylab='Second biplot component',xlim=xlim,ylim=ylim,cex.lab=0.5)
arrows(x0=0,y0=0,x1=Vtildes[,1],y1=Vtildes[,2],col='red', code=2, length=0.1)
text(.9*Vtildes[,1],Vtildes[,2],paste(colnames(Xssub)),pos=c(1,3),col='red',cex=0.5)
```



## Biplot
```{r}
nn <- paste(rownames(Xssub))
xlim <- ylim <- range(c(range(Vtildes),range(1.5*Vtildes)))
plot(Utilde,xlab='First biplot component',ylab='Second biplot component',xlim=xlim,ylim=ylim,cex.lab=1.25)
text(Utildes[,1:2],nn,pos=c(4,1,2),cex=0.5)
arrows(x0=0,y0=0,x1=Vtildes[,1],y1=Vtildes[,2],col='red', code=2, length=0.1)
text(.9*Vtildes[,1],Vtildes[,2],paste(colnames(Xssub)),pos=c(1,3),col='red',cex=0.5)

```



 
## Clustering

Distances
```{r}
d_euclidean <- dist(Xs)
d_manhattan <- dist(Xs, method="manhattan")

# S^(-1/2)
Sinvsqrt = V%*%diag(1/sqrt(l))

# X in mahalanobis space
Xinvsqrt = Xs%*%Sinvsqrt

d_mahalanobis <- as.dist(dist(Xinvsqrt))
```
Hierarchical Clustering
Eucledian
```{r}
hc <- hclust(d_euclidean)
plot(hc ,hang = -1, main="Euclidean dist - Complete linkage", cex=0.5)
```
Manhattan
```{r}
hc <- hclust(d_manhattan)
plot(hc ,hang = -1, main="Manhatta dist - Complete linkage", cex=0.5)
```



## k-means
```{r}
pc <- prcomp(Xs)$x
km1 <- kmeans(Xs, centers = 4 )
km2 <- kmeans(Xs, centers = 5 )

clus1 <- km1$cluster
clus2 <- km2$cluster

plot(pc[,1],pc[,2],col=clus1,pch=clus1,xlab='PC1',ylab='PC2',main="Plot with k-mean with k=4")
text(pc[,1]-0.1,pc[,2]-0.1, labels=rownames(X), cex=0.6)
plot(pc[,1],pc[,2],col=clus2,pch=clus2,xlab='PC1',ylab='PC2',main="Plot with k-mean with k=5")
text(pc[,1]-0.1,pc[,2]-0.1, labels=rownames(X), cex=0.6)

```



## Choosing k formally
```{r}
wss = c(sum(Xs^2))

for (i in 2:10) {
  km  = kmeans(Xs, centers = i)
  wss = c(wss,sum(km$withinss))
}
plot(wss,xlab="k",ylab="wss",main="WSS-Euclidean")
```




## With the subset

Distances
```{r}
d_euclideans <- dist(Xssub)
d_manhattans <- dist(Xssub, method="manhattan")

# S^(-1/2)
Sinvsqrts = Vs%*%diag(1/sqrt(ls))

# X in mahalanobis space
Xinvsqrts = Xssub%*%Sinvsqrts

d_mahalanobiss <- as.dist(dist(Xinvsqrts))
```


Hierarchical Clustering
Eucledian
```{r}
hcs <- hclust(d_euclideans)
plot(hcs ,hang = -1, main="Euclidean dist - Complete linkage", cex=0.5)
```
Manhattan
```{r}
hcs <- hclust(d_manhattans)
plot(hcs ,hang = -1, main="Manhatta dist - Complete linkage", cex=0.5)
```



## k-means
```{r}
pcs <- prcomp(Xssub)$x
km1s <- kmeans(Xssub, centers = 4 )
km2s <- kmeans(Xssub, centers = 5 )

clus1s <- km1s$cluster
clus2s <- km2s$cluster

plot(pcs[,1],pcs[,2],col=clus1s,pch=clus1s,xlab='PC1',ylab='PC2',main="Plot with k-mean with k=4")
text(pcs[,1]-0.1,pcs[,2]-0.1, labels=rownames(X), cex=0.6)
plot(pcs[,1],pcs[,2],col=clus2s,pch=clus2s,xlab='PC1',ylab='PC2',main="Plot with k-mean with k=5")
text(pcs[,1]-0.1,pcs[,2]-0.1, labels=rownames(X), cex=0.6)

```



## Choosing k formally
```{r}
wsssub = c(sum(Xssub^2))

for (i in 2:10) {
  kms  = kmeans(Xssub, centers = i)
  wsssub = c(wsssub,sum(km$withinss))
}
plot(wsssub,xlab="k",ylab="wss",main="WSS-Euclidean")
```



 
## Further thoughts

- Analysis with emissions and preferences was not fruitful although a nice idea

- Possible Solution: Perform Factor Analysis as

- Further explorations - Emissions and Taxes: Patterns underlying emission volumes and CO2 taxes on the country level + analysis with coded climate pledges to Paris Protocol

- Idea: Different groups acting according to stated behaviour and ones that do not + heterogenous dynamics according to actually being a high/low emitting country in comparison