---
title: "R Coding Sample"
author: "Jonas Wallstein"
date: "2022-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### The goal of this code is to find similiarities in emission profiles of several countries by performing a Principal Component Analysis on emission, GDP and climate pledge data


````{r Setup}
# Setup
library(readxl)
library(reshape)

data_path = paste(dirname(getwd()), "/data/", sep = "")
````

#### 1. Create a dataset on sector-level emissions from data from the climate trace project: https://climatetrace.org/
````{r}
filename_emissions = paste(data_path, "climatetrace_emissions_by_subsector_timeseries_interval_year_since_2020_to_2020.csv", sep = "")

emissions = read.csv(filename_emissions,sep = ",", header = TRUE)

# Aggregating Emissions over each Country
total_emissions = aggregate(emissions[,1], by = list(emissions$country, emissions$country_full), FUN = sum, na.rm = TRUE)
# Renaming variables
colnames(total_emissions) = c("Country Code","Emissions Data Country Name", "Emissions")

# Emissions per sector
per_sector_emissions = aggregate(emissions[,1], by = list(emissions$country, emissions$country_full, emissions$sector), FUN = sum, na.rm = T)
colnames(per_sector_emissions) = c("country code","country name", "sector", "emissions")
per_sector_emissions = reshape(per_sector_emissions, idvar = c("country code", "country name") , timevar = "sector", direction = "wide")
per_sector_emissions$emissions.total = total_emissions$Emissions


head(per_sector_emissions)
````

#### 2. Adding 2020 GDP data from https://data.worldbank.org/indicator/NY.GDP.MKTP.CD?end=2021&start=1960

````{r}
filename_gdp = paste(data_path, "GDP_worldbank.xls", sep = "")
# Reading data, Renaming variables to the correct names in the third row, Dropping the first three irrelevant rows
gdp = read_excel(filename_gdp, skip = 3, col_names = TRUE)

# Rename the 2020 GDP variable simply to GDP and Country code to match emission data
names(gdp)[names(gdp) == "2020"] <- "GDP"
names(gdp)[names(gdp) == "Country Code"] <- "country code"
# Drop all irrelevant variables
gdp = gdp[c("country code", "GDP")]
gdp$GDP = as.numeric(gdp$GDP)

df_per_sector = merge(per_sector_emissions, gdp, by = "country code", all = F)

# Saving dataframe as csv file
filename_per_sector_gdp = paste(data_path, "/output/per_sector_emissions_gdp.csv", sep = "")
write.csv(df_per_sector, file = filename_per_sector_gdp)

```

#### 3. Merging data on Climate Pledges from https://ourworldindata.org/grapher/net-zero-targets?country=SOM~BRA~MDG
````{r}
filename_pledges = paste(data_path, "/pledges/net-zero-targets.csv", sep = "")
pledges = read.csv(filename_pledges,sep = ",", header = TRUE)
names(pledges)[names(pledges) == "Code"] <- "country code"
names(pledges)[names(pledges) == "Year"] <- "net_zero_target"
pledges = pledges[c("country code", "net_zero_target")]

df_pledges = merge(df_per_sector, pledges, by = "country code", all = F)
rownames(df_pledges) = df_pledges[,1]
head(df_pledges)

# Saving dataframe as csv file
filename_merged_pledges= paste(data_path, "/output/pledges_emissions_gdp.csv", sep = "")
write.csv(df_pledges, file = filename_merged_pledges)
````
#### 4. Performing the Principle Component Analyis manually for practice
````{r}
# Transforming dataframe into matrix to perform matrix operations
X = as.matrix(df_pledges)
# Getting rid of non numercal variables
rownames(X) = X[,1]
X = X[,c(-1,-2)]
X = apply(X, 2, as.numeric)

# Scaling data
Xs = scale(X)
# Creating correlation matrix R
R = cor(Xs)
# Eigenvectors and values
E = eigen(R)$vectors
l = eigen(R)$values

R2 = l / sum(l) * 100
# The first two principal components explain 77% of the variance

# Perform the PCA
Y = Xs %*% E
colnames(Y) <- paste0("PC", 1:ncol(Y)) 
W = cor(X,Y)

````

#### 5. Plotting Variables and Observations in a PCA Plot
```{r}
# Limits for the graph with the variables
inferior = min(W[,1:2])
superior = max(W[,1:2])
limits = 1.01 * c(inferior, superior) + 0.1

# Defining the limits and the titles for the graph
plot(W[,1:2],
  type="n", 
  asp = 1, 
  main=paste("Variable PCA Plot, explained Variance is ",
  round(R2[1]+R2[2],1), "%"),
  xlim = limits, 
  ylim = limits, 
  xlab=paste("PC1(", round(R2[1],1), "%)"), 
  ylab=paste("PC2(", round(R2[2],1), "%)") 
)

# The axis
abline(v=0, col="blue")
abline(h=0, col="blue")

# Drawing the arrows
arrows(0,0,W[,1],W[,2])

# Adding the labels
text(1.1 * W[,1:2], colnames(X), col= "cadetblue4",cex=0.9)

`````

- The first principal component is largely determined by GDP and general emissions (without maritime emissions)

- The second principal component is mostly determined by the net-zero-target and by maritime emissions



```{r}
# Plotting observations
plot(Y[,1],Y[,2],xlab='PC1',ylab='PC2', type = "n", main=paste("Observation PCA Plot"))
points(Y[,1:2], pch="") 
text(Y[,1]-0.1,Y[,2]-0.1, labels=rownames(df_pledges), col = "cadetblue4")
`````

#### 6. Performing K-Means Clustering
```{r}
# Chose k = 4 clusters
km <- kmeans(Xs, centers = 4)
clus <- km$cluster
# Plotting Clusters
plot(Y[,1],Y[,2],col=clus,pch=clus,xlab='PC1',ylab='PC2',main="K-means clustering with 4 Clusters")
text(Y[,1]-0.1,Y[,2]-0.1, labels=rownames(X))

````

- China and the US form a cluster in most iterations and Indonesia, Brazil, Russia and Australia form another cluster
- To gain a better understanding of emission profiles, more data must be added





